---
title: "reticulate FotD"
output:
  bookdown::html_document2: default
---


<!-- <style type="text/css"> -->
<!-- .main-container { -->
<!--   max-width: 70% !important; -->
<!--   margin: auto; -->
<!-- } -->
<!-- </style> -->



```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(reticulate)
```


# Reticulate

you can use any conda environment to run python code in R.

```{r conda}
reticulate::conda_list(conda = "auto") %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
py_config()
```

```{r change, eval=FALSE, include=FALSE}
reticulate::conda_list(conda = "auto")
reticulate::use_condaenv(condaenv = 'p27', required = TRUE)
reticulate::py_config()
reticulate::use_condaenv(condaenv = 'r-reticulate', required = TRUE)
.rs.restartR()
reticulate::use_condaenv(condaenv = 'r-reticulate', required = TRUE)
reticulate::py_config()
```

# Radius Neighbors Classifier

Using epigenomic markers near insertion sites in CART theraphy to alucidate which sites
might predict treatment outcome.

```{r load}
df <- readRDS(file.path(.data_d,'tsne-df.rds'))
head(df) %>%
  kbl() %>%
  kable_paper() %>%
  scroll_box(width = "80%", height = "200px")
```

Show porjection of inserion sites into a latent space
```{r single}
ggplot(df, aes(x=X, y=Y, color=Response_class)) +
  geom_point()
```

```{r dual}
ggplot(df, aes(x=X, y=Y, color=Response_class)) +
  geom_point() +
  facet_wrap( ~ Response_class)
```

Formatting projection and responce of each site
```{r format}
pyx <- df %>% select(c(X,Y))
pyy <- df %>% select(Response_class)
```

# Running python code

The __"r"__ object __python__ contains an interface to the __R__ variables

```{python radius}
import pandas
from sklearn.neighbors import RadiusNeighborsClassifier

model = RadiusNeighborsClassifier(radius=0.1)
__ = model.fit(r.pyx,r.pyy.values.ravel())
yhat = model.predict(r.pyx)
```

The __"py"__ object in __R__ contains an interface to the __python__ variables

```{r yhats}
df_imp1 <- df[py$yhat == df$Response_class,]

 ggplot(df_imp1, aes(x=X, y=Y,color=Response_class)) +
 geom_point() +
 facet_wrap( ~ Response_class)
```

More information can be found in <https://rstudio.github.io/reticulate/>